Lightweight Real-Time Feature Monitoring

### 2020-07-13 paper.tex
Changed paper template

%%%% Proceedings format for most of ACM conferences (with the exceptions listed below) and all ICPS volumes.
\documentclass[sigconf]{acmart}
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[group-separator={,}]{siunitx}

%%%% As of March 2017, [siggraph] is no longer used. Please use sigconf (above) for SIGGRAPH conferences.

%%%% Proceedings format for SIGPLAN conferences 
% \documentclass[sigplan, anonymous, review]{acmart}

%%%% Proceedings format for SIGCHI conferences
% \documentclass[sigchi, review]{acmart}

%%%% To use the SIGCHI extended abstract template, please visit
% https://www.overleaf.com/read/zzzfqvkmrfzn

%
% defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

% Rights management information. 
% This information is sent to you when you complete the rights form.
% These commands have SAMPLE values in them; it is your responsibility as an author to replace
% the commands and values with those provided to you when you complete the rights form.
%
% These commands are for a PROCEEDINGS abstract or paper.
\copyrightyear{}
\acmYear{}
\setcopyright{none}
\acmConference[KDD-ADF-2019]{2nd KDD Workshop on Anomaly Detection in Finance}{August 5, 2019}{Anchorage, Alaska, USA}
\acmBooktitle{}
\acmPrice{}
\acmDOI{}
\acmISBN{}
\acmISBN{}

%
% These commands are for a JOURNAL article.
%\setcopyright{none}
%\acmJournal{TOG}
%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
%\acmDOI{10.1145/1122445.1122456}

@@

%
% The "author" command and its associated commands are used to define the authors and their affiliations.
% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
% used to denote shared contribution to the research.

\author[Conde]{João Conde}
\email{joao.conde@feedzai.com}
\affiliation{%
  \institution{Feedzai}
}

@@

\author[Ribeiro]{Pedro Manuel Pinto Ribeiro}
\email{pribeiro@dcc.fc.up.pt}
\affiliation{%
  \institution{FCUP}
}

\author[Restivo]{André Monteiro de Oliveira Restivo}
\email{arestivo@fe.up.pt}
\affiliation{%
	\institution{FEUP}
}

%
% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
% other information printed in the page headers. This command allows the author to define a more concise list
% of authors' names for this purpose.

%
% The abstract is a short summary of the work to be presented in the article.
\begin{abstract}
% The problem
Many real-time stream monitoring systems are static once deployed in a production environment. The engineers of those systems configure them under the assumption that future data flowing through the system roughly follows the same distribution as previously seen data. They do so consciously, to the best of their knowledge and available tools, keeping in mind that in the future they may need to reconfigure the system. Thus, even though the initial configuration of the system may be one of the best fits, over time, due to data pattern shifts, the initially deployed static system's performance gradually deteriorates. Data pattern shift detection refers to the process of finding patterns in data that do not conform to expected or usual behavior. Accurate and timely detection of data pattern deviations allows for immediate measures to be taken. Thus, the problem at hand is to determine when to reconfigure the system, \textit{e.g.}, a Machine Learning model, based on an analysis of the drifts in the stream of data.

%Approach/Solution and %Main contributions
In this thesis, we design a method that makes use of lightweight streaming aggregations and distribution divergence functions to alert about deviations in data patterns in real-time, while in the presence of large volumes of high velocity, highly skewed and seasonal data. We develop a two-phased and two-windowed method: first, we perform a batch analysis on a reference window and then a stream analysis over a sliding streaming window. We aggregate the contents of both windows using an approximate histogram aggregation based on Exponential Moving Averages (EMAs). We do this for each event and each of its fields, also denominated features in our context. We measure the divergence between both reference and sliding window EMA-based histograms for each feature with the Jensen–Shannon Divergence function, but other distance functions could be used. For each feature, we find out the probability value of each divergence measure and apply the Holm-Bonferroni multiple test correction to find out which probability values are lowest and their associated features. We generate alerts based on a user-defined probability threshold, for each feature.

We evaluate our method through a series of tests, some with synthetic datasets and some with real data. We further split the tests into single and multi-feature analysis. Our method accurately detected the introduced anomalies in the experiments using synthetic datasets while maintaining high throughput, for single and multi-feature analysis. However, experiments with real data were not as accurate. Despite not knowing the specific reasons for that, we defer further investigation to future work and formulate a set of hypotheses that might explain it and hence are worthy of pursuing next. Our set of experiments supports the claim that sliding window aggregations and distribution divergence methods can be combined to detect data pattern shifts in streaming scenarios with constant time and memory complexity.
\end{abstract}

%
% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

%\ccsdesc[500]{Computer systems organization~Embedded systems}
%\ccsdesc[300]{Computer systems organization~Redundancy}
%\ccsdesc{Computer systems organization~Robotics}
%\ccsdesc[100]{Networks~Network reliability}

@@

\section{Introduction}
Nice Intro

\section{Related Work}
\label{sec:RelatedWork}
Nice related work

\section{Method}
\label{sec:Method}

Such method wow

@@

%
% The next two lines define the bibliography style to be used, and the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{myrefs.bib}

>>>>

% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}
\usepackage{listings}
\usepackage{dirtytalk}
\usepackage[shortlabels]{enumitem}
\usepackage{amssymb}
\usepackage{minitoc}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{mathtools}

% pseudocode
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

%
% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

% Rights management information. 
% This information is sent to you when you complete the rights form.
% These commands have SAMPLE values in them; it is your responsibility as an author to replace
% the commands and values with those provided to you when you complete the rights form.
%
% These commands are for a PROCEEDINGS abstract or paper.
\copyrightyear{2020}
\acmYear{2020}
\setcopyright{acmlicensed}
\acmConference[MileTS '20]{MileTS '20: 6th KDD Workshop on Mining and Learning from Time Series}{August 24th, 2020}{San Diego, California, USA}
\acmBooktitle{MileTS '20: 6th KDD Workshop on Mining and Learning from Time Series, August 24th, 2020, San Diego, Claifornia, USA}
\acmPrice{15.00}
\acmDOI{10.1145/1122445.1122456}
\acmISBN{978-1-4503-9999-9/18/06}

%
% These commands are for a JOURNAL article.
%\setcopyright{acmcopyright}
%\acmJournal{TOG}
%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
%\acmDOI{10.1145/1122445.1122456}

@@

%
% The "author" command and its associated commands are used to define the authors and their affiliations.
% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
% used to denote shared contribution to the research.
\author[Conde]{João Conde}
\email{joao.conde@feedzai.com}
\affiliation{%
  \institution{Feedzai}
}

@@

\author[Ribeiro]{Pedro Ribeiro}
\email{pribeiro@dcc.fc.up.pt}
\affiliation{%
  \institution{University of Porto}
}

\author[Restivo]{André Restivo}
\email{arestivo@fe.up.pt}
\affiliation{%
	\institution{University of Porto}
}

%
% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
% other information printed in the page headers. This command allows the author to define a more concise list
% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%
% The abstract is a short summary of the work to be presented in the article.
\begin{abstract}
Many real-time stream monitoring systems are static once deployed in a production environment. Over time, due to data pattern shifts, the initially deployed static system's performance gradually deteriorates. Data pattern shift detection refers to the process of finding patterns in data that do not conform to expected or usual behavior. Accurate and timely detection of data pattern deviations allows for immediate measures to be taken. Thus, the problem at hand is to determine when to reconfigure the system, \textit{e.g.}, a Machine Learning model, based on an analysis of the drifts in the stream of data. In this paper, we present a method that makes use of lightweight streaming aggregations and distribution divergence functions to alert about deviations in data patterns in real-time. We generate alerts based on a user-defined probability threshold, for each of the event's fields, also denominated features in our context. We evaluate our method through a series of tests, some with synthetic datasets and some with real data. We further split the tests into single and multi-feature analysis. Our method accurately detected the introduced anomalies in the experiments using synthetic datasets while maintaining high throughput, for single and multi-feature analysis. However, experiments with real data were not as accurate. Despite not knowing the specific reasons for that, we defer further investigation to future work and formulate a set of hypotheses that might explain it and hence are worthy of pursuing next. Our set of experiments supports the claim that sliding window aggregations and distribution divergence methods can be combined to detect data pattern shifts in streaming scenarios with constant time and memory complexity.
\end{abstract}

%
% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10002951.10003227.10003351.10003446</concept_id>
       <concept_desc>Information systems~Data stream mining</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10010520.10010570.10010574</concept_id>
       <concept_desc>Computer systems organization~Real-time system architecture</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Data stream mining}
\ccsdesc[500]{Computer systems organization~Real-time system architecture}

@@

\section{Introduction}
In the past decade, applications have become increasingly data-driven, placing data at the center of application design. For different use cases, the data is processed in different ways for different purposes. For instance, e-commerce platforms such as Alibaba\footnote{https://alibaba.com} need to process a large number of daily transactions while ensuring that sales run smoothly and that products are delivered to customers' homes. Streaming services, like YouTube\footnote{https://youtube.com} and Netflix\footnote{https://netflix.com}, try to guarantee that media content reaches up to millions of users simultaneously. Social networks are responsible for generating large volumes of data. Twitter\footnote{https://twitter.com}, for example, generates more than 500 million \textit{tweets} (posts in Twitter) per day\footnote{https://blog.twitter.com/engineering/en\_us/a/2013/new-tweets-per-second-record-and-how.html}. \textit{Tweets} may contain text, media content or both. Cybersecurity applications are yet another example of time-sensitive, data-driven applications. The entities that work in these use cases need to monitor user accesses and their actions in the network, where timely detection of intruders is critical to prevent them from tampering with the underlying system.

All of these applications generate large volumes of data which lead to the creation of multiple real-time unbounded datasets, also known as data streams. The information that flows through such a stream can be analyzed on-the-fly or stored for later processing. The former is best known as stream processing, whereas the latter is known as batch processing. Streaming data is known for being non-stationary \cite{Gama-Knowledge-Discovery} where the value of the produced information lies in its recency \cite{Kolajo-Big-data-stream-SLR}. \textit{Recency} is measured under different scales for different use cases. For instance, consider the monitoring of geological data to forecast possible natural disasters such as earthquakes versus the monitoring of a computer network for intruder detection. In the former case, after the forecast of a future earthquake, governmental authorities need a couple of days to launch an evacuation plan and keep the population safe. However, in the intruder detection scenario, the decision of removing access to a user must be done as soon as possible, preferably in a few seconds. Hence, while in the first scenario information retrieved within a day would still be recent, in the second scenario, information is considered recent and relevant if delivered within seconds (or even milliseconds).

Consider the use case where we want to detect credit card fraud on the data stream of all the transactions made on Amazon\footnote{https://amazon.com} by every US citizen on a single day. In this case, Amazon's transaction stream will need to be monitored by another system, a fraud detection one. Such a system would have to process all incoming transactions and decide, in a fraction of a second, if a transaction is fraudulent or not. Unlike the fraud detection system, our system is not mission-critical. Hence, it must monitor the data stream looking for data pattern deviations in real-time but with a low memory footprint.

To summarize, in this paper we design a lightweight real-time system that monitors a stream of high volumes of high velocity, highly skewed and seasonal data and detects pattern shifts relative to a reference period. The motivation to detect data stream pattern deviations in real-time is to know when to reconfigure the mission-critical system (\textit{e.g.}, retrain the Machine Learning model) before its performance decays. We want to build a lightweight enough solution that can be integrated into existing workflows since having a real-time system monitoring other real-time systems consuming as many resources as the latter would be far too costly.

\section{Related Work}
\label{sec:RelatedWork}

\section{Method}
\label{sec:Method}
A data stream is a continuous collection of events, each timestamped and with multiple fields. We denote each of the fields as features. Our goal is to detect data pattern deviations for each feature in a streaming fashion. Such shifts will be measured between a static in time reference period versus the observed reality during the streaming phase. Figure \ref{fig:timelines} shows the separated reference and streaming timelines.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/timeperiods.png}
      \caption[]{Reference and streaming timelines}
      \label{fig:timelines}
    \end{center}
\end{figure}
To detect data pattern shifts, we do not need to directly compare the entire window contents of the reference and sliding windows. This approach would require storing all window contents hence making the memory consumption grow linearly with target sliding window and/or reference period size. Instead, we want to use a \textit{(a)} fixed cost in-memory aggregation that can be \textit{(b)} incrementally maintained during the streaming phase. Additionally, we want a streaming aggregation that \textit{(c)} reliably encodes the distribution of each feature.

One aggregation that encodes distributions properly are histograms. Histograms are a representation of the distribution of numerical data thus being ideal to encode the distribution of feature values, satisfying \textit{(c)}. Histograms require storing a fixed-size list of bins and corresponding counts per bin regardless of the volume of data, satisfying property \textit{(a)}. Adding or removing an element to the histogram aggregation is tantamount to determining the correct bin and incrementing or decrementing the associated count, respectively, thus satisfying property \textit{(b)}. Recall Section \ref{sec:back-swag-algs} where we discussed generic sliding window aggregation algorithms and remember the Subtract-On-Evict (SOE) algorithm (Algorithm \ref{pseudo:soe}). Computing a histogram aggregation over a sliding window of size $W$ using SOE implies storing all $W$ elements in memory to know the element to remove after a new one is inserted. This violates our constant in-memory property \textit{(a)} which is why we chose to use the Exponential Moving Averages (EMAs) analyzed in Section \ref{sec:emas}. EMA-like histograms computed over sliding windows do not require the storage of window contents. An EMA-like histogram is \textit{(a)} constant in memory, \textit{(b)} maintained in real-time and \textit{(c)} encodes the distribution of each feature values. These EMA-like histograms are further explained in Section \ref{sec:ema-hist}.

Since the reference window is static, \textit{i.e.}, it is not a sliding window, we can use exact histogram aggregations. Hence, for each feature, we build an exact reference histogram aggregation. On the other hand, in streaming, we use one EMA-like approximate histogram aggregation per feature (Figure \ref{fig:timelines-hists}).

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/timelines-hists.png}
      \caption[]{Reference and Streaming timelines and corresponding histogram aggregations}
      \label{fig:timelines-hists}
    \end{center}
\end{figure}

To compare each feature's reference and target sliding window histogram aggregation we need a measure of divergence. We chose to use the Jensen–Shannon Divergence (JSD) \cite{JSD} but other statistical tests are possible such as Kolmogorov-Smirnov or Wasserstein \cite{EncyclopaediaMathematics}. The JSD takes as input two histograms and outputs a value between 0 and 1. The closer this value is to 0 the more similar the distributions are. On the other hand, a JSD value of 1 represents two different distributions. We chose to use the JSD metric due to its simplicity and because it was used in previous work successfully \cite{SAMM}, leaving as future work other possible tests.

Given a divergence or distance value for the two aggregations (reference and target) we need a measure of normal statistical fluctuations. Since we consider the reference period as the normal state we can find the normal distribution for the divergence by sampling smaller windows of data from the reference period in a batch analysis. This allows us to define a threshold based on percentiles, which gives an estimate for the probability of a certain value occurring, rather than using a constant hand-picked threshold for the divergence value (details in Section \ref{sec:sampling-batch}).

To implement the proposed methodology, we devised a two-phase method. The first phase relies on a batch analysis that, for each feature, builds a reference histogram and computes the distribution of divergence values (details presented in Section \ref{sec:batch-phase}). The second phase (detailed in Section \ref{sec:stream-phase}) focuses on efficiently updating the target sliding window histogram for each feature and periodically performing a divergence test between the static reference aggregation and the target sliding one, alerting for changes if need be. The proposed method currently works only for numerical features, that is, only for event fields or variables that are of numeric type, but can be extended to other types.

\subsection{Exponential Moving Average Histogram} \label{sec:ema-hist}
As stated, we compute histogram aggregations for the reference and target sliding windows. The computation of the reference histogram is done once in batch. Since the reference data is fixed and does not change over time, computing this histogram in batch is more efficient and easier to implement in a distributed system like Spark. In addition, a batch computation also allows us to compute an exact reference histogram. However, the target sliding window histogram is an EMA-like histogram to ensure that we are able to encode the feature distribution with constant space complexity and maintain it in real-time.

In Section \ref{sec:emas} we discussed Exponential Moving Averages (EMAs) and how they compute sliding window aggregations without actually storing the window contents. Definition \ref{def:tuple-ema} shows how to compute an EMA using its recursive formula. An EMA state or value depends only on the previous one multiplied by an exponential decay. Recall that a tuple-based decay is given by:
\[ 2^{- \frac{1}{n_{1/2}}} \]
where $n_{1/2}$ is commonly denoted as a tuple-based half-life.

Algorithm \ref{pseudo:ema-hist} shows how we update our EMA histogram. We assume we have access to global \textit{counts} and \textit{bins} arrays and use a suppression factor of $2^{- \frac{1}{n_{1/2}}}$ with $n_{1/2}$ as the tuple-based half-life. 
The histogram aggregation is updated for each new streaming value \texttt{val} using the \texttt{processValue(val)} function, that increments the correct histogram bin (note that the \textit{bins} array defines the bin boundaries while the \textit{counts} array actually holds each bin's count) and applies the EMA suppression to every bin.
\begin{algorithm}[!htb]
    \caption[EMA histogram]{EMA histogram sliding window emulation}
    \label{pseudo:ema-hist}
    \begin{algorithmic}[1]
        \Function{getBin}{val}
            \State \Return{binarySearch(bins, val)}
        \EndFunction

        \Function{processValue}{val}
            \State $bin \gets getBin(val)$
            \State $counts[bin] \gets counts[bin] + 1$ \Comment{// Increment correct bin}

            \ForEach {$count \in counts$} \Comment{// Apply EMA suppression to all bins}
                \State $count \gets count \times 2^{- \frac{1}{n_{1/2}}}$
            \EndFor
        \EndFunction
    \end{algorithmic}
\end{algorithm}

After processing \textit{x} tuples the oldest tuple is \emph{smoothed} by a factor of $2^{- \frac{x}{n_{1/2}}}$. If we process $x = 4n_{1/2}$ tuples the oldest tuple would have its contribution reduced by
\[ 2^{- \dfrac{4n_{1/2}}{n_{1/2}}} = 2^{-4} \]
which is approximately 0.06. In other words, the \textit{x-th} event would have 6\% contribution towards the EMA state or value meaning older events would have even less weight.

We choose to emulate windows with four times the half-life size (or $4n_{1/2}$), meaning we discard events whose contribution is less than 6\%. We set the half-life to be one-quarter of the size of the window we want to emulate. The size of the window we want to monitor varies from situation to situation, but in the limited scope of this thesis we set to monitor a period extending up to one month (at an error level of about 6\%) so we set the half-life to be the number of tuples that correspond to one week worth of data. This will give more weight to the most recent week but it will include contributions decaying to 6\% in the fourth oldest week.

\subsection{Batch Analysis Phase} \label{sec:batch-phase}
The goal of the batch phase will be to build our reference aggregation for each feature and get to know the distribution of divergence or distance values between the reference and EMA-like aggregations. In other words, we:

\begin{enumerate}
    \item Obtain a reference histogram for each feature using all data in the reference period

    \item Compute the distribution of distances between the reference histogram and random windows of data and corresponding histograms
\end{enumerate}

This distribution of values can then later be compared to a given value of divergence for a single sample to find out how probable that value is to occur (if it is very low an alert may be raised).

\subsubsection*{Building the Reference Histogram for each Feature}

First, for each feature, we compute the reference histogram from the reference period dataset. This is an exact histogram aggregation. When building it we ensure the histogram has equal counts (or equal height) for all bins which often results in different sized bins. We make no assumption on the dataset distributions and hence use equal height histograms because they adjust better to wildly varying ones. Equal height means we have more bins covering very dense regions (regions with many data points) and fewer bins in lower density regions.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.3]{figures/ref-hist.png}
      \caption[]{Reference equal-height histogram}
      \label{fig:ref-hist}
    \end{center}
\end{figure}

\subsubsection*{Finding the Distribution of Distances for each Feature} \label{sec:sampling-batch}

Secondly, we aim to build the histogram that encodes the distribution of expected distance values, to later on threshold the observed distance values during the online phase. To that end, we make $S$ samples of transactions (in Section \ref{sec:nsamples} we discuss the minimum number $S$ of samples to make), each with the same tuple-based size of the target sliding window we will use in streaming. Each sample is a contiguous block of transactions, thus preserving the order of transactions and the time-dependency property of a time-series.

For each sample, and each feature, we compute the approximated histogram using the bins computed for the reference histogram (Figure \ref{fig:ref-hist-bins}). We also add two extra bins that will cover the region to the left and the right of our histogram, respectively. In other words, we add a bin that covers the region between $-\infty$ and our first bin and another that covers values from the last bin to $+\infty$. This way, observed values outside of the reference period will be placed in these special bins, either left or right.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.3]{figures/ref-bins.png}
      \caption[]{Reference equal-height histogram}
      \label{fig:ref-hist-bins}
    \end{center}
\end{figure}

Each of the histograms is an EMA-like histogram. An EMA-like histogram is essentially a collection of EMA-counts, as described in Section \ref{sec:ema-hist}. Each sample is processed event by event and a discount factor is applied per event to each bin (EMA expiration factor), thus building an approximated histogram aggregation, using the reference bins, for each feature of that sample. This procedure is illustrated in Figure \ref{fig:EMA-hist-build}.
\begin{figure}[!htb]
    \begin{center}
    %\hspace{2cm}
      \includegraphics[scale=0.8]{figures/ema-hist.png}
      \caption[]{Reference equal-height histogram}
      \label{fig:EMA-hist-build}
    \end{center}
\end{figure}
We compute multiple sample histograms to encode distributions over smaller time periods that may exist within the large reference period. We use an approximated histogram to mimic the target approximated histogram we will incrementally maintain in the streaming environment.

For each feature, we have now \textit{S} histograms, one for each sample. For each of the \textit{S} histograms, we compute the distance between it and the reference histogram. Figure \ref{fig:compute-sample-distances} illustrates this process as we apply our distance function for each sample histogram and the reference histogram, obtaining \textit{S} distance measurements.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/compute-sample-distances.png}
      \caption[Compute sample's distance values]{Compute as many distance values as samples, one for each sample-reference histogram pair}
      \label{fig:compute-sample-distances}
    \end{center}
\end{figure}

For each feature, we end up with \textit{S} distance values. These $S$ measurements are distance measurements between random samples of data and the reference period. Hence, we claim we now know the distribution of expected distance values. Given a new distance value measured between another random window and the reference window, we can compute its probability and produce alerts if the probability is below a certain threshold. Note that the histograms for each sample are EMA-like histograms (as detailed in Section \ref{sec:ema-hist}) just like the target sliding window histogram. This ensures a certain degree of fidelity in the test we make in streaming because we measure the divergence between a random sample of data (our streaming histogram or any of the sample's histogram) and the reference one.

It is important to understand why we want to know the distribution of divergence values instead of just thresholding the divergence value itself. For instance, given that JSD values vary between 0 and 1, why not just set to alert if the measured JSD value is 0.7? The reason is that a divergence value may fall on different percentiles for different features. In other words, the same \textit{i-th} percentile will correspond to a different divergence value for different features. Consider the following left and right-skewed distance distributions represented by Figures \ref{fig:skewed-left-distro} and \ref{fig:skewed-right-distro}, respectively.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.2]{figures/skewed-left-distro.png}
      \caption[]{Left skewed distribution of distance values}
      \label{fig:skewed-left-distro}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.2]{figures/skewed-right-distro.png}
      \caption[]{Right skewed distribution of distance values}
      \label{fig:skewed-right-distro}
    \end{center}
\end{figure}

Let's assume that the histogram from Figure \ref{fig:skewed-left-distro} encodes the distribution of expected distance values for a feature \textit{x1} and Figure \ref{fig:skewed-right-distro} does the same but for a feature \textit{x2}. Setting a user-defined threshold of \textit{$\alpha$=0.6} for both features would yield very different results. For feature \textit{x1} we would be reporting distance values above the 100th-percentile which would not be the case for feature \textit{x2}. Instead, we define our threshold as a percentile of the distribution and not as a hard-coded distance constant. For instance, in this case, we would define the threshold to be the \textit{i}-th percentile, which would correspond to a threshold value of 0.4 for feature \textit{x1} and 0.6 for feature \textit{x2}.

Hard-coding a threshold value to be used for all features is cumbersome, impractical and error-prone. The distance threshold must be computed based on the distribution of distances.

\subsubsection*{Burn-In Period at System Initialization}
When we boot up our system in the streaming phase we initially have empty histogram aggregations. The period of time where you discard the alerts produced by the system until it processes enough data to produce accurate reports is commonly denoted as the burn-in period. To avoid this burn-in period we propose that the target sliding window approximate histogram aggregation for each feature is initialized using the last sample's histogram for that feature.

\subsubsection*{Batch Phase Artifacts} \label{sec:batch-artifacts-summary}
By the end of the batch phase, we should have, for each feature: 
\begin{itemize}
    \item a reference histogram
    \item a list of distance values that represent its distribution
    \item the last sample's histogram, to be used as burn-in period for the target sliding window histogram
\end{itemize}

\subsection{Streaming Phase} \label{sec:stream-phase}
In the streaming or online phase, the system will process event by event in a true sliding window fashion (recall Section \ref{sec:windows}) and perform a change detection test periodically. This phase makes use of all of the batch phase artifacts, namely, for each feature, the reference histogram, the distribution of distance values and the last sample's histogram.

For each feature, we initialize the approximate histogram with the last sample's data, from the batch phase. Then, for each incoming event and each of the event's features, we update our streaming approximate histogram for that feature by adding the new value and then "cutting a slice" from the top of the histogram bars to simulate eviction, as described in Section \ref{sec:ema-hist}. This is all we do, regarding event processing.

In order to obtain alerts, we periodically perform a multi-feature change detection test. The test is done by first computing the distance between the reference histogram for a feature \textit{x} and the target sliding window histogram of the same feature. This distance value is then tested against the known distribution of distance values for feature \textit{x}. In other words, we test to see if the percentile of the computed distance is in the distribution of distance values and alert if above a certain percentile, \textit{e.g.}, the 99th-percentile. When we perform the test, for each feature \textit{x}, we compute the percentile of the distance measured between the reference and target sliding window histogram aggregations. The probability value or \textit{p-value} of a distance value is:

\[\textit{p-value\textsubscript{distance\textsubscript{x}}} = 1 - percentile(\textit{distance\textsubscript{x}})\]

For example, if we define alerts to have 1\% or less probability to happen, it means the system will raise an alert for distance values above the 99th-percentile or \textit{p-value} = 0.01. Generally speaking, we alert a feature \textit{x} as divergent at a given timestamp if for a given probability threshold $\alpha$:
\[ \textit{percentile(distance\textsubscript{x})} > 1 - \alpha \]
or:
\begin{equation}
    \label{eq:alert-test}
    \textit{p-value\textsubscript{distance\textsubscript{x}}} < \alpha
\end{equation}

\subsubsection*{Multiple Test Correction} \label{sec:multi-test}
Note that we periodically apply this test (Equation \ref{eq:alert-test}) for each feature, effectively repeating the analysis. When considering several hypotheses, the problem of multiplicity arises: the more hypotheses are checked, the higher the probability of obtaining false positives. In other words, when repeating a test multiple times the multiple test or simultaneous statistical inference problem \cite{MultiTestProblem-Pubmed, MultiTestProblem-Dickhaus2014, MultiTestProblem-Miller1966, multiple-test-correction-geoffrey} arises which states that the more inferences are made, the more likely erroneous inferences are to occur. In the literature, many multiple test correction methods are proposed, but we opted to use the Holm-Bonferroni multiple test correction because it is a simple and easy to compute correction that can reduce the number of false negatives while having as target to control the number of false positives. Furthermore, it does not require to assume any independence between the tested hypothesis (as other methods do) which would almost surely be a wrong assumption for realistic datasets. Therefore we leave as future work the testing of other correction methods. In statistics, Family-Wise Error Rate (FWER) \cite{MultivariateMT, MultitestTamhane2018AdvancesIP} is the probability of having one or more false positives when performing multiple hypotheses tests, \textit{i.e.}, considering one hypothesis true when it is not. The Holm-Bonferroni method \cite{HolmBonferroni} is one of many approaches for controlling the Family-Wise Error Rate \textit{(FWER)} by adjusting the rejection criteria for each hypothesis.

\subsubsection*{Holm-Bonferroni Correction} \label{sec:holmbonferroni}
The Holm-Bonferroni correction takes Equation \ref{eq:alert-test} and corrects the right-side threshold $\alpha$ for each feature. This correction is done by dividing the threshold $\alpha$ by a correction factor, as described next.

First, we compute the \textit{p-value} of each feature. We then order our features by \textit{p-value}, in ascending order. Next, starting from the first feature, the one with the smallest \textit{p-value}, and proceeding until the end of the list of features, we check if:
\begin{equation}
    p-value_{distance_{x}} < \alpha / (m - k)
    \label{eq:corrected-pvalue}
\end{equation}
with $\alpha$ as the probability to reject the null hypothesis, \textit{m} as the number of hypotheses to test, in our case the number of features and corresponding \textit{p-values} and \textit{k} as the 0-based index of the current feature. We stop at the first feature that fails this test, and every feature processed up until this one is considered to be a divergent feature and alerts are raised.

\subsection{Final Remarks}

\subsubsection*{Minimum Number of Samples} \label{sec:nsamples}
In Section \ref{sec:sampling-batch} we stated we made \textit{S} samples in the batch phase. But what is the optimal number of samples to make? How can we mathematically derive a formula for the number of samples to make and avoid \textit{ad hoc} user-defined constants?

When checking for alerts we are performing one test for each feature, hence we perform \textit{n\textsubscript{features}} tests. We apply a multiple test correction on the tests performed and want to set the FWER to $\alpha$. In other words, we want to alert deviations for the full set of tests globally that have probability $\alpha$ < 1\%, for example.

We use the Holm-Bonferroni correction and, for each feature, we need to reliably estimate the upper tail of the distribution which corresponds to the \textit{$\alpha$ / n\textsubscript{features}} region. Each sample we do translates into a distance value and the collection of these data points builds the distribution of distance values. We want to be sure that the number of samples \textit{S} that we do generates enough distance points to have a high probability of having one or more points in the upper tail region. To that end, we define the probability of not having at least one point in the upper tail of the distance distribution as $\gamma$.

In \cite{SAMM}, the authors prove that the minimum number of samples $S$ needed to satisfy these probability requirements is:
\begin{equation}
    \label{eq:optimal-n-samples}
    \textit{S} = \ceil[\bigg]{\frac{\log\gamma}{log(1 - \frac{\alpha}{n_{features}})}}
\end{equation}

\subsubsection*{Method classification}
Similarly to the methods presented in Sections \ref{kifer} and \ref{sec:SAMM}, we use a two-windowed model that can take uni or multivariate input. However, we perform a multivariate analysis of all feature's time-series with the use of the Holm-Bonferroni correction. Hence, we say our method is multivariate. We also alert entire windows or subsequences as anomalous. Thus, according to the taxonomy presented in Section \ref{sec:outliers}, our method is a dissimilarity-based multivariate subsequence outlier detection method.

\subsubsection*{Time complexity of the streaming phase}
The critical phase of our method is the stream analysis because it has to process new events and alert for possible deviations in real-time.

For each event and for each feature, we update our EMA histogram using Algorithm \ref{pseudo:ema-hist} which has to perform a binary search to find the correct bin to insert our new value and traverses all bins to apply a suppression factor. Using $n$ as the number of bins, we have a time complexity of $O(log(n) + n)$, for each feature. However, since $n$ is a fixed and small constant, \textit{i.e.}, more or less 100 bins, we say that this algorithm has constant time complexity. In other words, our time complexity does not depend on the size of the emulated streaming windows.

\subsubsection*{Space complexity of the streaming phase}
We store $n$ bins per feature, so we have a space complexity of $O(n)$ per feature. Again, since $n$ is a small and fixed constant, we claim to have constant space complexity for ever-growing volumes of data.

@@

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``acmart'' document class provide powerful taxonomic tools for you to help readers find your work in an online search.

The ACM Computing Classification System --- \url{https://www.acm.org/publications/class-2012} --- is a set of classifiers and concepts that describe the computing discipline. Authors can select entries from this classification system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases of the authors' choosing, providing a more flexible way of describing the research being presented.
